Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2019-12-20 article-information-2019.tex

\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$. In Friedman's texts, a Gaussian noise term was added to create a continuous output variable for testing spline regression methodologies. In this manuscript, the signal generating function and input features are modified in several ways.  Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings. 
%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so. For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.  Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing. The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows. The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.  Within the training set, a 5 fold cross validation indicator is used for training and assessing all models. For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.

@@

\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the percent of the control class that receives a favorable outcome.  Another long-standing measure of DI is SMD. SMD is typically used to assess disparities in continuous variables, such as income differences in employment analyses, or interest rate differences in lending. It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}. The SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.

@@

\noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.  Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of variables, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.  In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.  Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.

@@

Results for constrained models on the simulated data are displayed in Subsections \ref{ssec: c_v_uc_sim} -- \ref{ssec:dis_sim}. {Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models. \textcolor{red}{...}

@@

Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. $g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results. Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Accuracy is measured at the best F1 threshold for each model.

@@

@@

Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the numeric value of the reported values can interpreted as the absolute mean impact of each $X_j$ on $g^\text{MGBM}$ in the margin space.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
\label{fig:sim_mgbm_glob}
\caption{Global mean Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data. Tree SHAP values are reported in the margin space, prior to the application of the logit link function.}
\end{figure}

% highlight monotonicity 
PD and ICE are always displayed with a histogram herein to highlight any sparse regions in an input feature's domain. Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction. Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
\label{fig:sim_mgbm_glob_pdp_ice}
\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/sim_xnn_glob.png}
\label{fig:sim_xnn_glob}
\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the simulated test data.}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
\label{fig:sim_xnn_glob_ridge}
\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the simulated test data.}
\end{figure}

@@

Deep SHAP values are reported in the probability space, after the application of the logit link function.

@@

11.1\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.  Of note is that 11.1\% is not a meaningful difference without context.  If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 11.1\% could represent an extremely large difference and would require remediation.  But if they represent substantially different populations, then 11.1\% could represent a reasonable deviation from parity.  As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 11.1\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population. Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.  It is also highly statistically significant (not shown), which together would indicate that there may be evidence of illegal DI.  As with ME and other measures, the reasonableness of this disparity is not clear outside of context.  However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.

@@

Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data. This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g. numerical encoding of categorical features and missing values. This preprocessing appears to hamstring some of the tree-based models' inherent capabilities. $g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$.

@@

Figure \ref{fig:mort_mgbm_glob_pdp_ice} also displays PD and ICE for the unconstrained feature property value. Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns established by ICE curves. While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$, and lower,  exhibit different prediction behavior with respect to property value. Some individuals at these lower percentiles display monotonically decreasing prediction behavior while others appear to show fluctuating prediction behavior. Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn. For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions. However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value. This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci}. However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.

@@

\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_k$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function. For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income. In combination, Figure \ref{fig:mort_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.

@@

The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own. However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g. the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}. High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}. This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{17}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems. Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena. Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML. However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks and that explanations can mask ML discrimination. Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory. Accuracy, discrimination, and security characteristics of a system can change over time as well. Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems. Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.

>>>>

\noindent where $\mathbf{X}_{\text{Friedman},j}$ are random uniform features in $[0,1]$. In Friedman's texts, a Gaussian noise term was added to create a continuous output feature for testing spline regression methodologies. In this manuscript, the signal generating function and input features are modified in several ways.  Two binary features, a categorical feature with five discrete levels, and a bias term are introduced into $f$ to add a degree of complexity that may more closely mimic real-world settings. 
%When generating the data, random draws for each observation are constructed such that the features can be correlated with each other, some highly so. For example, X1 and X3 have a Pearson correlation coefficient equal to 0.58.
For binary classification analysis, the Gaussian noise term is replaced with noise drawn from a logistic distribution and coefficients are re-scaled to be one fifth of the size of those used by Friedman, and any $f(\mathbf{X})$ value above 0 is classified as a positive outcome, while $f(\mathbf{X})$ values less than or equal to zero are designated as negative outcomes.  Finally, $f$ is augmented with two hypothetical protected class-control features with known dependencies on the binary outcome to allow for discrimination testing. The simulated data is generated to have eight input features, twelve after numeric encoding of categorical features, and a binary outcome, two class-control features, and 100,000 rows. The simulated data is then split into a training and test set, with 80,000 and 20,000 observations, respectively.  Within the training set, a 5 fold cross validation indicator is used for training and assessing all models. For an exact specification of the simulated data, see the software resources referenced in Subsection \ref{ssec:soft}.

@@

\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the percent of the control class that receives a favorable outcome.  Another long-standing measure of DI is SMD. SMD is typically used to assess disparities in continuous features, such as income differences in employment analyses, or interest rate differences in lending. It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}. The SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.

@@

\noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.  Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of features, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.  In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.  Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.

@@

Model fit is roughly uniform for $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. Given that little or no trade-off is required in terms of model to fit to use the constrained models, intrinsic interpretability, post-hoc explainability, and discrimination are explored further for the $g^\text{MGBM}$ and $g^\text{XNN}$ models in Sections \ref{ssec:int_ex_sim} - \ref{ssec:dis_sim}.

@@

Table \ref{tab:sim_acc} presents a variety of fit metrics for the $g^\text{GBM}$, $g^\text{MGBM}$, $g^\text{ANN}$, and $g^\text{XNN}$ on the simulated test data. $g^\text{GBM}$ exhibits the best performance, but all models give relatively similar fit results.

@@

\noindent Interpretability and explainability benefits of the constrained models appear to come at little cost to overall model performance, or in the case of $g^\text{ANN}$ and $g^\text{XNN}$, no cost at all. $g^\text{XNN}$ actually shows slightly better fit than $g^\text{ANN}$ across accuracy, area under the curve (AUC), logloss, and root mean squared error (RMSE). Accuracy is measured at the best F1 threshold for each model.

@@

Global mean absolute Shapley value feature importance for $g^{\text{MGBM}}(\mathbf{X})$ on the simulated test data is displayed in Figure \ref{fig:sim_mgbm_glob}. Tree SHAP values are reported in the margin space, prior to the application of the logit link function, and the reported numeric values can be interpreted as the mean absolute impact of each $X_j$ on $g^\text{MGBM}(\mathbf{X})$ in the simulated test data in the $g^\text{MGBM}$ margin space.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/sim_mgbm_glob.png}
\caption{Global mean absolute Tree SHAP feature importance for $g^\text{MGBM}$ on the simulated test data.}
\label{fig:sim_mgbm_glob}
\end{figure}

\noindent As expected, the $X_{\text{Friedman},j}$ features from the original \citet{friedman1979tree} and \citet{friedman1991multivariate} formula are the main drivers of $g^\text{MGBM}(\mathbf{X})$ predictions, with encoded versions of the augmented categorical and binary features contributing less on average to $g^\text{MGBM}(\mathbf{X})$ predictions.

Figure \ref{fig:sim_mgbm_glob_pdp_ice} highlights PD, ICE, and histograms of the most important features from Figure \ref{fig:sim_mgbm_glob}. PD gives an indication of the estimated average prediction of $g^\text{MGBM}(\mathbf{X})$ in the test data for some $X_j$. ICE serves several purposes. It indicates the trustworthiness of the PD in the presence of interactions and correlation, it can be an interaction detector, and is itself a type of sensitivity analysis where certain $X_j$ are varied in single records under $g^\text{MGBM}$ to provide a glimpse at local model behavior. PD and ICE are always displayed with a histogram herein to highlight any sparse regions in an input feature's domain. Because most ML models will always issue a prediction on any datum with a correct schema, it's crucial to consider whether a given model learned enough about an observation to make an accurate prediction. Viewing PD and ICE along with a histogram is a convenient method to visually assess whether a prediction is reasonable and based on sufficient training data.

\begin{figure}[H]
\centering
\includegraphics[width=15cm]{img/sim_mgbm_glob_pdp_ice.png}
\caption{PD, ICE for 10 observations across selected percentiles of $g^\text{MGBM}(\mathbf{X})$, and histograms for the three most important input features of $g^\text{MGBM}$ on the simulated test data.}
\label{fig:sim_mgbm_glob_pdp_ice}
\end{figure}

$X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ were positively monotonically constrained under $g^\text{MGBM}$ for the simulated data, and positive monotonicity looks to be confirmed on average with PD and at numerous local percentiles of $g^\text{MGBM}(\mathbf{X})$ with ICE. Also, as the PD curves generally follows the patterns of the ICE curves, PD is likely an accurate representation of average feature behavior for $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$. Since PD and ICE curves do not obviously diverge, $g^\text{MGBM}$ is likely not modeling strong interactions, despite the fact that known interactions are included in the simulated data signal generating function in Equation \ref{eq:sim}. The one-dimensional monotonic constraints probably hinder the modeling of such interactions, but do not strongly affect overall $g^\text{MGBM}$ accuracy, perhaps due to noise in the simulated data. This is an interesting result for responsible ML practitioners. In some noisy scenarios, monotonicity constraints can increase model interpretability without causing a drastic drop in model accuracy, even when known interactions and non-monotonic behavior exist in training data.

Local Shapley values for records at the 10\textsuperscript{th}, 50\textsuperscript{th}, and 90\textsuperscript{th} percentiles of $g^\text{MGBM}(\mathbf{X})$ in the simulated test data are displayed in Figure \ref{fig:sim_mgbm_loc}. These Tree SHAP values are reported in the margin space of $g^\text{MGBM}$, and each $\phi_j$ value in Figure \ref{fig:sim_mgbm_glob_pdp_ice} represents the difference in $g^\text{MGBM}(\mathbf{x^{(i)}})$ and the average of $g^\text{MGBM}(\mathbf{X})$ associated with some input feature $X_j$ \cite{molnar}. Accordingly, the logit of the sum of the Shapley values and the Shapley intercept will be the $g^\text{MGBM}(\mathbf{x})$ prediction in the probability space, for any $\mathbf{x}$.

@@

\noindent The Shapley values in Figure \ref{fig:sim_mgbm_loc} appear to be a logical result. For the lower prediction at the 10\textsuperscript{th} percentile of $g^\text{MGBM}(\mathbf{X})$, the largest local contributions are negative and the majority of local contributions are also negative. At the median of $g^\text{MGBM}(\mathbf{X})$, local contributions are roughly split between positive and negative values, and at the 90\textsuperscript{th} of $g^\text{MGBM}(\mathbf{X})$, most large contributions are positive. In each case, large local contributions follow global importance results in Figure \ref{fig:sim_mgbm_glob} as well.

Figure \ref{fig:sim_xnn_glob} shows global mean absolute Shapley feature importance for $g^\text{XNN}$ on the simulated test data, using the approximate Deep SHAP technique. Deep SHAP values are reported in the probability space, after the application of the logit link function. They are also calculated from the projection layer of $g^\text{XNN}$. Thus, the Deep SHAP values in Figure \ref{fig:sim_xnn_glob} are the estimated average absolute impact of each $X_j$ in the projection layer and probability space of $g^\text{XNN}$ for the simulated test data.

\begin{figure}[H]
\centering
\includegraphics[width=8cm]{img/sim_xnn_glob.png}
\caption{Global mean Deep SHAP feature importance for $g^\text{XNN}$ on the simulated test data.}
\label{fig:sim_xnn_glob}
\end{figure}

\noindent Like $g^\text{MGBM}$, $g^\text{XNN}$ ranks the $X_{\text{Friedman},j}$ features higher in terms of importance than the categorical and binary features. The consistency between the feature rankings of $g^\text{MGBM}$ and $g^\text{XNN}$ is somewhat striking, given their different hypothesis families and architectures. Both $g^\text{MGBM}$ and $g^\text{XNN}$ rank $X_{\text{Friedman},1}$, $X_{\text{Friedman},2}$, and $X_{\text{Friedman},4}$ as the most important features, both place $X_{\text{Categorical},2}$ and $X_{\text{Categorical},3}$ above the $X_{\text{Binary},1}$ and $X_{\text{Binary},2}$ features, both rank $X_{\text{Binary},1}$ above $X_{\text{Binary},2}$, and both place the least importance on $X_{\text{Categorical},4}$ and $X_{\text{Categorical},0}$.

@@

\begin{figure}[H]
\centering
\includegraphics[width=17cm]{img/sim_xnn_glob_ridge.png}
\caption{\textbf{A.} Output layer $\gamma_k$ weights, \textbf{B.} corresponding $n_k$ ridge functions, and \textbf{C.} associated projection layer $\beta_k$ weights for $g^\text{XNN}$ on the simulated test data.}
\label{fig:sim_xnn_glob_ridge}
\end{figure}

In combination, Figure \ref{fig:sim_xnn_glob_ridge} \textbf{A}, \textbf{B}, and \textbf{C} help practitioners understand which original input $X_j$ features are weighed heavily in each $n_k$ subnetwork, and which $n_k$ subnetworks have a strong influence on $g^\text{XNN}(\mathbf{X})$.

@@

@@

11.1\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.  Of note is that 11.1\% is not a meaningful difference without context.  If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 11.1\% could represent an extremely large difference and would require remediation.  But if they represent substantially different populations, then 11.1\% could represent a reasonable deviation from parity.  As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 11.1\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population. Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.  It is also highly statistically significant (not shown, but available in resources discussed in Section \ref{sec:soft}), which together would indicate that there may be evidence of illegal DI.  As with ME and other measures, the reasonableness of this disparity is not clear outside of context.  However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.

@@

Table \ref{tab:mort_acc} shows that $g^\text{ANN}$ and $g^\text{XNN}$ noticeably outperform $g^\text{GBM}$ and $g^\text{MGBM}$ on the mortgage data. This is at least partially due to the preprocessing required to present directly comparable post-hoc explainability results and to use neural networks and tensorflow, e.g. numerical encoding of categorical features and missing values. This preprocessing appears to hamstring some of the tree-based models' inherent capabilities. $g^\text{GBM}$ models trained on non-encoded data with missing values repeatedly produced AUC values of $\sim0.81$ (not shown, but available in resources discussed in Section \ref{sec:soft}).

@@

Figure \ref{fig:mort_mgbm_glob_pdp_ice} also displays PD and ICE for the unconstrained feature property value. Unlike DTI ratio and LTV ratio, PD for property value does not always follow the patterns established by ICE curves. While PD shows monotonically increasing prediction behavior on average, apparently influenced by large predictions at extreme $g^{\text{MGBM}}(\mathbf{X})$ percentiles, ICE curves for individuals at the 40\textsuperscript{th} percentile of $g^{\text{MGBM}}(\mathbf{X})$, and lower,  exhibit different prediction behavior with respect to property value. Some individuals at these lower percentiles display monotonically decreasing prediction behavior while others appear to show fluctuating prediction behavior. Property value is strongly right-skewed, with little data regarding high-value property from which $g^{\text{MGBM}}$ can learn. For the most part, reasonable predictions do appear to be carried from more densely populated regions to more sparsely populated regions. However, prediction fluctuations at lower $g^{\text{MGBM}}(\mathbf{X})$ percentiles are visible, and in a sparse region of property value. This divergence of PD and ICE can be indicative of an interaction affecting property value under $g^{\text{MGBM}}$ \cite{ice_plots}, and analysis by surrogate decision tree did show evidence of numerous potential interactions in lower predictions ranges of $g^{\text{MGBM}}(\mathbf{X})$ \cite{art_and_sci} (not shown, but available in resources discussed in Section \ref{sec:soft}). However, fluctuations in ICE can also be caused by overfitting or leakage of strong non-monotonic signal from important constrained features into the modeled behavior of non-constrained features.

@@

\noindent \ref{fig:mort_xnn_glob_ridge} \textbf{C} displays the sparse $\beta_k$ weights of the projection layer $\sum_j\beta_{k,j}x_j$ hidden units that are associated with each $n_k$ subnetwork ridge function. For instance, subnetwork $n_3$ is influenced by large weights for LTV ratio, no introductory rate period flag, and introductory rate period, whereas subnetwork $n_9$ is nearly completely dominated by the weight for income.

@@

The black-box nature of ML, the perpetuation or exacerbation of discrimination by ML, or the security vulnerabilities inherent in ML are each serious and difficult problems on their own. However, evidence is mounting that these harms can also manifest as complex intersectional challenges, e.g. the \textit{fairwashing} or \textit{scaffolding} of biased models with ML explanations, the privacy harms of ML explanations, or the adversarial poisoning of ML models to become discriminatory \cite{shokri2019privacy}, \cite{fair_washing}, \cite{scaffolding}.\footnote{See: \href{https://www.theguardian.com/technology/2016/mar/24/tay-microsofts-ai-chatbot-gets-a-crash-course-in-racism-from-twitter}{\textit{Tay, Microsoft's AI chatbot, gets a crash course in racism from Twitter}}.}\textsuperscript{,}\footnote{While the focus of this paper is not ML security, proposed best-practices from that field do point to transparency of ML systems as a mitigating factor for some ML attacks and hacks \cite{papernot2018marauder}. High system complexity is sometimes considered a mitigating influence as well \cite{hoare19811980}. This is sometimes known as the \textit{transparency paradox} in data privacy and security, and it likely applies to ML security as well, especially in the context of interpretable ML models and post-hoc explanation.\textsuperscript{\textcolor{red}{24}}}\textsuperscript{,}\footnote{See: \href{https://hbr.org/2019/12/the-ai-transparency-paradox}{\textit{The AI Transparency Paradox}}.} Again, this text makes no claims that the opacity, discrimination, or security problems in ML have been solved, even treated as independent problems. Instead, this text aims to highlight these issues as both singular entities and non-static intersectional phenomena. Practitioners should of course consider the discussed interpretable modeling, post-hoc explanation, and discrimination testing approaches as at least partial remedies to the black-box and discrimination issues in ML. However, they should also consider that explanations can ease model stealing, data extraction, and membership inference attacks and that explanations can mask ML discrimination. Additionally, high-stakes, human-centered, or regulated ML systems should generally be built and tested with robustness to adversarial attacks as a primary design consideration, and specifically to prevent ML models from being poisoned or otherwise altered to become discriminatory. Accuracy, discrimination, and security characteristics of a system can change over time as well. Simply testing for these problems at training time, as presented in Sections \ref{sec:m_and_m} and \ref{sec:res}, is not adequate for high-stakes, human-centered, or regulated ML systems. Accuracy, discrimination, and security should be monitored in real-time and over time, as long as a model is deployed.
